---
title: "Untitled"
author: "Yun Zhao"
date: "12/2/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
#Load
library(stringr)
library(utils)
library(readr)
library(here)
library(googledrive)

#Wrangle
library(janitor)
library(tidyverse)
library(car)
library(crosstalk)
library(RSQLite)
library(dplyr)

#Vector
library(rgdal)
library(sf)
library(sp)
library(fs)
library(rgeos)
library(geojson)
library(geojsonio)

#Raster
library(raster)
library(fpc)
library(rasterVis)
library(terra)
library(RStoolbox)
library(htmlwidgets)

#Graphs
library(plotly) # graphs
library(ggplot2)
library(GGally)

#Maps
library(maptools)
library(tmap)
library(tmaptools)
library(GISTools)
library(OpenStreetMap)
library(RColorBrewer)
library(rgeos)
library(rmapshaper)
library(leafpop)
library(leaflet)
#library(mapview)

#Stats
library(broom) # or tidymodels
library(tidypredict)
library(car)
library(rstatix)
library(spatstat)
library(spdep)

```

When the American Home Owners' Loan Corporation (HOLC) classified the "credit-worthiness" of neighbourhoods, it discriminated areas based on race. To this day, this redlining still affects access to outdoor space, clean air, trees, local temperature, and deaths due to weather. 
"Vegetation has a large impact on local temperature, with heat being the leading cause of weather related deaths in the USA. As a result you have been enlisted as a consultant and tasked to conduct an analysis exploring temperature differences across the HOLC "graded neighbourhoods" in Los Angeles."

### Project scope: bullet points okay. show awareness of data
Load and stack raster
Calcuate LST from raster with NDVI
Join raster and holc
raster::extract raster with holc
transform to utf
veg value 0.3?????
map: variation in temp
Summary statistics: histogram of temp

### Load LSAT rasters
```{r cache=TRUE}
#o<-drive_download(" url ",
                  #path="data/LC08_L1TP_041036_20210503_20210508_02_T1.tar",
                  #overwrite=T)

# Unzip tar.gz
listfiles<-dir_info(here::here("data", "landsat"))%>%
  dplyr::filter(str_detect(path, ".tar"))%>%  #subset row
  dplyr::select(path)%>%  #subset column
  dplyr::pull()%>%   #pull is $
  print()%>%
  as.character()%>% # returns text representation of a raster or extent object for quick gen
  utils::untar(exdir=here::here("data", "landsat"))

#try this here
#raster::compareRaster(listfiles$LC08_L1TP_203023_20190513_20190521_01_T1_B1,
 #             listfiles$LC08_L1TP_203023_20190513_20190521_01_T1_B8)
```

### Stack LSAT rasters
```{r}
# List your raster files except band 8
listlandsat<-dir_info(here::here("data", "landsat"))%>%
  dplyr::filter(str_detect(path, "[B123456790].TIF")) %>%   
  dplyr::select(path)%>%
  pull()%>%
  as.character()%>%
  stack()

# get band 8 and save as raster
b8list<-dir_info(here::here("data", "landsat"))%>%
  dplyr::filter(str_detect(path, "[B8].TIF")) %>%
  dplyr::select(path)%>%
  pull()%>%
  as.character()%>%
  raster()

# resampling: ngb is a nearest neighbour sampling method
b8correct <- b8list%>%
  #resample(raster to be resampled, raster with parameters to sample to)
  raster::resample(., listlandsat$LC08_L1TP_041036_20210503_20210508_02_T1_B1, 
             method = "ngb") %>%
  # Write out the raster to a file. str_c joins into single vector
  raster::writeRaster(.,str_c(here::here("data", "landsat"), 
                  names(b8list), 
                  sep="/"),
            format='GTiff', 
            overwrite=TRUE)

# Make resampled band 8 into a raster
b8backin<-dir_info(here::here("data", "landsat"))%>%
  dplyr::filter(str_detect(path, "[B8].tif")) %>%
  dplyr::select(path)%>%
  pull()%>%
  as.character()%>%
  raster()

# Add resampled band 8 into the stack
listlandsat <- listlandsat %>%
  addLayer(., b8backin)

# Check if resampling succeeded (use this earlier too)
raster::compareRaster(listlandsat$LC08_L1TP_041036_20210503_20210508_02_T1_B1,
              listlandsat$LC08_L1TP_041036_20210503_20210508_02_T1_B8)

```

### Stack v.2
```{r}
# List your raster files except band 8
listlandsat<-dir_info(here::here("data", "landsat"))%>%
  dplyr::filter(str_detect(path, "[B1234567890].TIF")) %>%   
  dplyr::select(path)%>%
  pull()%>%
  as.character()%>%
  stack()
```

### Clip raster to LA County study area
```{r}
tracts <- st_read(here::here("data", 
                                    "Census Tracts 2010",
                                    "geo_export_aa2e3d74-5f12-4539-ad39-ab773bb7501c.shp"))

holc <- st_read(here::here("data", "fullshpfile", "shapefile", "holc_ad_data.shp"))


# check they have the same Coordinate Reference System (CRS)
st_crs(tracts)$proj4string
st_crs(holc)$proj4string
st_crs(listlandsat)$proj4string

tracts <- tracts %>%
  st_transform(.,st_crs(listlandsat)$proj4string)
holc <- tracts %>%
  st_transform(.,st_crs(listlandsat)$proj4string)
```

### Clip raster to LA County study area
```{r cache=TRUE}
# Crop and mask
lsatmask <- listlandsat%>%
  raster::crop(., tracts) %>%
  raster::mask(., tracts)

# have raster names end in 'mask'
outputfilenames <-
  str_c(here::here("data/landsat_mask/"), names(lsatmask), "_mask" , sep="")

lsatmask %>%
  raster::writeRaster(., outputfilenames, 
              bylayer=TRUE, 
              format='GTiff', 
              overwrite=TRUE)
```

### Data Explore
```{r}
# Select and name Bands 1-7
la<-dir_info(here::here("data", "landsat_mask")) %>%
  dplyr::filter(str_detect(path, "[B1234567]_mask.tif")) %>%
  dplyr::filter(str_detect(path, "B11", negate=TRUE)) %>% #because picks up 11
  dplyr::select(path)%>%
  pull()%>%
  stack()
names(la) <- c('ultra-blue', 'blue', 'green', 'red', 'NIR', 'SWIR1', 'SWIR2')

# true colour composite
la_rgb <- stack(la$red, la$green, la$blue)
# false colour composite
la_false <- stack(la$NIR, la$red, la$green)

# Map true/false colors
la_rgb %>%
  raster::plotRGB(.,axes=TRUE, stretch="lin")
la_false %>%
  raster::plotRGB(.,axes=TRUE, stretch="lin")
```

More Maps
```{r}
#set window size, plot single bands
par(mfrow = c(2,2))
#plot(la$blue, main = "Blue")
#plot(la$green, main = "Green")
plot(la$red, main = "Red")
plot(la$NIR, main = "NIR")
#plot(la$SWIR1, main = "SWIR1")#plot(manc$SWIR2, main = "SWIR2")

#WARNING
# Correlation, fancier with ggplot2, GGally, terra
pairs(la[[1:7]])
la %>%
  terra::as.data.frame(., na.rm=TRUE)%>%
  dplyr::sample_n(., 100)%>%
  ggpairs(.,axisLabels="none")
```


### Basic Raster Calculations: Vegetation
```{r}
# Define function to calculate NDVI
NDVIfun <- function(NIR, Red) {
  NDVI <- (NIR - Red) / (NIR + Red)
  return(NDVI)
}

ndvi <- NDVIfun(la$NIR, la$red)
  


# Map
ndvi %>%
  plot(.,col = rev(terrain.colors(10)), main = "Landsat-NDVI")

# Histogram
ndvi %>%
  hist(., breaks = 50000, main = "NDVI Histogram", xlim = c(-10,10))

# Reclassify based on histogram
veg <- ndvi %>%
  reclassify(., cbind(-Inf, 0.3, NA))

# Remap
veg %>%
  plot(.,main = 'Possible Veg cover')

# Map with context
manc_rgb %>%
  plotRGB(.,axes = TRUE, stretch = "lin", main = "Landsat True Color Composite")

veg %>%
  plot(., add=TRUE, legend=FALSE)
```

ndvi graph looks weird

## Advanced Raster Calculations

### Get Values from MTL file
```{r}
# RStoolbox MTL() option?

# Get greyscale and grescale from .MTL file
MTL<-dir_info(here::here("Rasters", "LSAT_data")) %>%
  dplyr::filter(str_detect(path, "MTL.txt")) %>%
  dplyr::select(path)%>%
  pull()%>%
  readMeta()

 #To see all the attributes
head(MTL)

# Extract values from MTL
offsetandgain <-MTL %>%
  getMeta("B10_dn", metaData = ., what = "CALRAD")
offsetandgain
#offsetandgain <- subset(MTL$CALRAD, rownames(MTL$CALRAD) == "B10_dn")
```

### Calculate Land Surface Temp with vegetation info
```{r}
summary(listlandsat)
# TOA, Use band 10 raster layer
TOA <- offsetandgain$gain *
  lsatmask$LC08_L1TP_203023_20190513_20190521_01_T1_B10 + 
  offsetandgain$offset

# Brightness Temp
Calidata <- MTL$CALBT%>%
  terra::as.data.frame()%>%
  mutate(Band=rownames(.))%>%
  filter(Band=="B10_dn")
# subset the columns
K1 <- Calidata %>%
  dplyr::select(K1)%>%
  pull()
K2 <- Calidata %>%
  dplyr::select(K2)%>%
  pull()
Brighttemp <- (K2 / log((K1 / TOA) + 1))

# fractional vegetation
facveg <- (ndvi-0.2/0.5-0.2)^2

#emissivity
emiss <- 0.004*facveg+0.986

# constants
Boltzmann <- 1.38*10e-23
Plank <- 6.626*10e-34
c <- 2.998*10e8
p <- Plank*(c/Boltzmann)

#define remaining variables
lambda <- 1.09e-5

#run the LST calculation Land Surface Temperature
LST <- Brighttemp/(1 +(lambda*Brighttemp/p)*log(emiss))
# check the values
LST

LST <- LST-273.15
plot(LST)
```

### Urban Area and Temp
```{r}
NDBI=((lsatmask$LC08_L1TP_203023_20190513_20190521_01_T1_B6-
         lsatmask$LC08_L1TP_203023_20190513_20190521_01_T1_B5)/
        (lsatmask$LC08_L1TP_203023_20190513_20190521_01_T1_B6+
        lsatmask$LC08_L1TP_203023_20190513_20190521_01_T1_B5))

NDBIfunexample <- NDVIfun(lsatmask$LC08_L1TP_203023_20190513_20190521_01_T1_B6,
                          lsatmask$LC08_L1TP_203023_20190513_20190521_01_T1_B5)

plot(values(NDBI), values(LST))

# stack the layers
computeddata <- LST%>%
  stack(.,NDBI)%>%
  terra::as.data.frame()%>%
  na.omit()%>%
  # take a random subset
  dplyr::sample_n(., 500)%>%
  dplyr::rename(Temp="layer.1", NDBI="layer.2")

# check the output
plot(computeddata$Temp, computeddata$NDBI)

heat<-ggplot(computeddata, aes(x = NDBI, y = Temp))+
  geom_point(alpha=2, colour = "#51A0D5")+
  labs(x = "Temperature", 
       y = "Urban index",
       title = "Manchester urban and temperature relationship")+
   geom_smooth(method='lm', se=FALSE)+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))

# interactive plot
ggplotly(heat)

# plotly, htmlwidgets
heat<-ggplot(computeddata, aes(x = NDBI, y = Temp))+
  geom_point(alpha=2, colour = "#51A0D5")+
  labs(x = "Temperature", 
       y = "Urban index",
       title = "Manchester urban and temperature relationship")+
   geom_smooth(method='lm', se=FALSE)+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))

# interactive plot
ggplotly(heat)

# plot whole data set
computeddatafull <- LST%>%
  stack(.,NDBI)%>%
  terra::as.data.frame()%>%
  na.omit()%>%
  # take a random subset
  dplyr::rename(Temp="layer.1", NDBI="layer.2")

hexbins <- ggplot(computeddatafull, 
                  aes(x=NDBI, y=Temp)) +
  geom_hex(bins=100, na.rm=TRUE) +
  labs(fill = "Count per bin")+
  geom_smooth(method='lm', se=FALSE, size=0.6)+
  theme_bw()

ggplotly(hexbins)
```

### Statistical Summary
```{r}
#library(rstatix)
Correlation <- computeddatafull %>%
  cor_test(Temp, NDBI, use = "complete.obs", method = c("pearson"))
Correlation

#critical t-value
# T is simply the calculated difference represented in units of standard error. The greater the magnitude of T, the greater the evidence against the null hypothesis. https://blog.minitab.com/en/statistics-and-quality-data-analysis/what-are-t-values-and-p-values-in-statistics
#2-sided test, degrees of freedom
abs(qt(0.05/2, 198268))

computeddatafull %>%
  pull(Temp)%>%
  length()

length(computeddatafull)
```

##  LSOA/MSOA stats
```{r}
# read in LSOA and project it to match Manchester boundary
UK_LSOA <- st_read(here::here("Shapefiles", 
                                          "Lower_Super_Output_Area_(LSOA)_IMD2019__(WGS84)-shp",
                                          "5878ebcf-f091-4bde-8864-aa547dd45afd2020330-1-8g4usn.8evuq.shp"))%>%
  st_transform(., 32630)

# read in MSOA and project it to match Manchester boundary
MSOA <- st_read(here::here("Shapefiles",
                           "MSOA_EngWal_Dec_2011_Generalised_ClippedEW_0",
                           "Middle_Layer_Super_Output_Areas_December_2011_Generalised_Clipped_Boundaries_in_England_and_Wales.shp")) %>%
  st_transform(., 32630)

qtm(manchester_boundary)
#select only MSOA within boundary
manchester_MSOA <- MSOA[manchester_boundary, , op=st_within]
qtm(manchester_MSOA)

#select only LSOA that intersect MSOA
manchester_LSOA <- UK_LSOA[manchester_MSOA,]
qtm(manchester_LSOA)
```
### Manchester Temp
```{r}
# extract mean LST value per LSOA
LST_per_LSOA <- raster::extract(LST, manchester_LSOA, fun=mean, na.rm=TRUE, df=TRUE)
#?raster::extract is like a spatial version of groupby + summarize

# add the LSOA ID back
LST_per_LSOA$FID<-manchester_LSOA$FID # so can later join

# join the average temp to the sf
manchester_LSOA_temp <- manchester_LSOA %>%
  left_join(.,
            LST_per_LSOA,
            by="FID")%>%
  dplyr::rename(temp=layer) # new name = old name
```

### Manchester Urban
```{r}
#define urban as NDBI greater than 0
NDBI_urban<- NDBI > 0

# Sum the pixels that are grater than 0 per LSOA
NDBI_urban_per_LSOA <- raster::extract(NDBI_urban, manchester_LSOA, na.rm=TRUE, df=TRUE, fun=sum)

# list the pixels per LSOA
NDBI_per_LSOA_cells <- raster::extract(NDBI_urban, manchester_LSOA, na.rm=TRUE, df=TRUE, cellnumbers=TRUE)

#count the pixels per LSOA
NDBI_per_LSOA2_cells<- NDBI_per_LSOA_cells %>%
  count(ID)

#add the LSOA ID to the urban area
NDBI_urban_per_LSOA$FID<-manchester_LSOA$FID

#add the LSOA ID to the number of cells
NDBI_per_LSOA2_cells$FID<-manchester_LSOA$FID

#join these two
Urban_info_LSOA <- NDBI_urban_per_LSOA %>%
  left_join(.,
            NDBI_per_LSOA2_cells,
            by="FID")

# remove what you don't need and rename
Urban_info_LSOA_core_needed <- Urban_info_LSOA %>%
  dplyr::rename(urban_count=layer, 
                LSOA_cells=n) %>%
  dplyr::select(urban_count,
         LSOA_cells,
         FID)%>%
  dplyr::mutate(percent_urban=urban_count/LSOA_cells*100)

# join the data 
# one sf with temp and % urban per LSOA
manchester_LSOA_temp_urban <- manchester_LSOA_temp %>%
  left_join(.,
             Urban_info_LSOA_core_needed,
             by="FID")
```

```{r}
Places <- st_read(here::here("prac7_data", 
                                          "greater-manchester-latest-free.shp",
                                          "gis_osm_places_free_1.shp")) %>%
   st_transform(., 32630)

manchester_Places <- Places[manchester_boundary,]%>%
  filter(fclass=="city")
```

```{r}
# this first bit makes the box bigger
# so we can have a north arrow not overlapping the data
# see: https://www.jla-data.net/eng/adjusting-bounding-box-of-a-tmap-map/
bbox_new <- st_bbox(manchester_LSOA_temp_urban) # current bounding box

yrange <- bbox_new$ymax - bbox_new$ymin # range of y values

bbox_new[4] <- bbox_new[4] + (0.1 * yrange) # ymax - top
bbox_new[2] <- bbox_new[2] - (0.1 * yrange) # ymin - bottom

# the plot starts here
library(tmap)
tmap_mode("plot")
# set the new bbox
# remove bbox=bbox_new to see the difference
tm1 <- tm_shape(manchester_LSOA_temp_urban, bbox = bbox_new) + 
  tm_polygons("temp",
              palette="OrRd",
              legend.hist=TRUE,
              title="Temperature")+
  tm_shape(manchester_Places, bbox=bbox_new)+
  tm_dots(size=0.1, col="white")+
  tm_text(text="name", size=0.75, ymod=-0.5, col="white", fontface = "bold")+
  #tm_legend(show=FALSE)+
  tm_layout(frame=FALSE,
            legend.outside=TRUE)+
  tm_compass(type = "arrow", size=1, position = c("left", "top")) +
  tm_scale_bar(position= c("left", "bottom"), breaks=c(0,2,4), text.size = .75)
  #tm_credits("(a)", position=c(0,0.85), size=1.5)

tm1

```


```{r}
library(biscale)
library(cowplot)
library(sysfonts)
library(extrafont) 
library(showtext) # more fonts
#font_add_google("Lato", regular.wt = 300, bold.wt = 700) # I like using Lato for data viz (and everything else...). Open sans is also great for web viewing.
showtext_auto()

# create classes
data <- bi_class(manchester_LSOA_temp_urban, x = temp, y = percent_urban, style = "jenks", dim = 3)

#ggplot map
map <- ggplot() +
 geom_sf(data = data, mapping = aes(fill = bi_class), color=NA, lwd = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "DkViolet", dim = 3) +
  geom_sf(data = manchester_MSOA, mapping = aes(fill=NA), color="black", alpha=0, show.legend = FALSE)+
  geom_sf(data=manchester_Places, mapping=aes(fill=NA), color="white", show.legend = FALSE)+
  geom_sf_text(data=manchester_Places, aes(label = name, hjust = 0.5, vjust = -0.5),
               nudge_x = 0, nudge_y = 0,
               fontface = "bold",
             color = "white",
             show.legend = FALSE,
             inherit.aes = TRUE)+
  labs(
    title = "",
    x="", y=""
  ) +
  bi_theme()

legend <- bi_legend(pal = "DkViolet",
                    dim = 3,
                    xlab = "Temperature ",
                    ylab = "% Urban",
                    size = 8)

credit<- ("Landsat dervied temperature and urban area, taken 13/5/19")

# combine map with legend
finalPlot <- ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, 0.1, 0.1, 0.2, 0.2)
  #draw_text(credit, 0.68, 0.1, 0.2, 0.2, size=10)
finalPlot
```



