---
title: "wk5_prac"
author: "Yun Zhao"
date: "11/3/2021"
output: html_document
---

# Notes

bedroom is at ward - convert to borough
could also probably just use Local Area District Data
Hotel is poly - summarise over borough
Airbnb is points - summarise over borough

```{r library, include = FALSE}
library(sf)
library(tidyverse)
library(janitor)
library(here)
library(tidyverse)
library(tmap)
library(grid)
```

# Read and Wrangle Data
st_join by geometry containing, rather than merge or left join by column names. 

### London Wards and Boroughs
https://data.london.gov.uk/dataset/statistical-gis-boundary-files-london 

```{r wards, include = FALSE}

Londonboroughs <- st_read(here::here("Shapefiles",
                                  "statistical-gis-boundaries-london",
                                  "ESRI",
                                  "London_Borough_Excluding_MHW.shp"))%>%
  clean_names()%>%
  st_transform(., 27700)
                          
Londonwards <- st_read(here::here("Shapefiles",
                              "statistical-gis-boundaries-london",
                              "ESRI",
                              "London_Ward_CityMerged.shp"))%>%
  clean_names()%>%
  st_transform(., 27700)

```

## Join to Wards

### NOMIS - Office for National Statistics website
https://www.nomisweb.co.uk/census/2011/data_finder
bedrooms per ward

```{r bed to borough, include = FALSE}

ward_beds <- read_csv(here::here("CSVs",
                                 "/nomis_bedrooms_bulk.csv"))%>%
  clean_names()

beds_borough <- Londonwards %>%
  left_join(.,
            ward_beds,
            by = c("gss_code" = "geography_code"))%>%
  group_by(lb_gss_cd, borough)%>%   
  summarise(total_beds_per_borough=sum(bedrooms_all_categories_number_of_bedrooms_measures_value,
                                       na.rm=TRUE))%>%
  st_drop_geometry()
  
```

### Airbnb detailed listings
http://insideairbnb.com/get-the-data.html

```{r airbnb, include=FALSE}
Airbnb <- read_csv(here::here("CSVs",
                   "Airbnb_Listings_London.csv"))%>%
  st_as_sf(., coords = c("longitude", "latitude"),
           crs = 4326)%>%
  clean_names()%>%
  st_transform(., 27700)
  #filter(room_type == 'Entire home/apt' & availability_365 =='365')
  
Airbnb_borough <- Airbnb %>%
  filter(room_type == 'Entire home/apt')%>%
  filter(availability_365 =='365')%>%
  st_join(Londonboroughs, .)%>%
  group_by(gss_code)%>%
  summarise(airbnbcount=n())%>%
  st_drop_geometry()
```

### OSM Geofabrik
http://download.geofabrik.de/europe/great-britain/england/greater-london.html

```{r osm_hotels include = FALSE}

OSM <- st_read(here::here("Shapefiles",
                          "greater-london-latest-free.shp",
                          "gis_osm_pois_a_free_1.shp")) %>%
                 clean_names()%>%
                 st_transform(., 27700)
                 #filter(fclass == 'hotel')
               
OSM_hotels <- OSM %>%
  filter(fclass == "hotel")%>%
  st_join(Londonboroughs, .)%>%
  group_by(gss_code)%>%  #needed ""
  summarise(hotelcount=n())%>%
  st_drop_geometry()
```

### World Cities, ESRI
https://hub.arcgis.com/datasets/6996f03a1b364dbab4008d99380370ed_0/explore

```{r extra_info, include = FALSE}

Worldcities <- st_read(here::here("Shapefiles", "World_Cities/World_Cities.shp"))

Worldcities_extract <- Worldcities %>%
  clean_names()%>%
  filter(cntry_name == "United Kingdom")%>%
  filter(city_name=="London" | city_name == "Birmingham" | city_name == "Edinburgh")
  # or symbol | works for certain operators such as filter
```

## London Outline, GADM 
https://gadm.org/download_country.html

```{r GADM, include=FALSE}  


UK_outline <- st_read(here::here("Shapefiles", 
                                 "gadm36_GBR_shp", 
                                 "gadm36_GBR_0.shp"))%>%
  st_transform(., 27700)

```

## Join all beds data

```{r all together, include = FALSE}

all_data <- Londonboroughs %>%
  left_join(.,
            beds_borough,
            by = c("gss_code" = "lb_gss_cd"))%>%
  left_join(.,
             Airbnb_borough,
            by = c("gss_code" = "gss_code"))%>%
  left_join(.,
             OSM_hotels,
            by = c("gss_code" = "gss_code"))

mapnorm <- all_data %>%
  mutate(airbnb_norm=airbnbcount/total_beds_per_borough)%>%
  mutate(hotelnorm=hotelcount/total_beds_per_borough)

```

# Mapping

## Static Map
```{r static map}

tmap_mode("plot")

#breaks = c(0, 5, 12, 26, 57, 286)
#t<-getJenksBreaks(mapnorm$airbnb_norm, k=5)

tm1<-tm_shape(mapnorm)+
  tm_polygons("airbnb_norm", style="jenks", palette="PuBu")+
  tm_legend(show=FALSE)+
  tm_layout(frame=FALSE)+
  tm_credits("(a) % Airbnb Beds", position=c(0,0.85), size=1.5)

tm2<-tm_shape(mapnorm)+
  tm_polygons("hotelnorm",
              style="jenks",
              palette="PuBu")+
  tm_legend(show=FALSE)+
  tm_layout(frame=FALSE)+
  tm_credits("(b) % Hotel Beds", position=c(0,0.85), size=1.5)

newbb <- c(xmin=-296000, ymin=5408, xmax=655696, ymax=1000000)

UK_outlinecrop <- UK_outline$geometry %>%
  st_crop(., newbb)

tm3<-tm_shape(UK_outlinecrop)+
    tm_polygons(col="darkslategray1")+
    tm_layout(frame=FALSE)+
    tm_shape(Worldcities_extract)+
    tm_symbols(col="red", scale=.5)+
    tm_text("city_name", xmod=-1, y=-0.5)

legend <- tm_shape(mapnorm)+
  tm_polygons("hotelnorm",
              style="jenks",
              palette="PuBu")+
  tm_scale_bar(position=c(0.2,0.06), text.size=0.6)+
  tm_compass(north=0, position=c(0.65,0.6))+
  tm_layout(legend.only = TRUE, legend.position=c(0.2,0.25), asp=0.1)+
  tm_credits("(c) OpenStreetMap contributors and Airbnb", 
             position=c(0.0,0.0))

t=tmap_arrange(tm1, tm2, tm3, legend, ncol=2)
#t
tmap_save(t, here::here('R_Maps/wk5_static_plot.png'))

# or arrange with grid package
#grid.newpage()
#pushViewport(viewport(layout=grid.layout(2,2)))
#print(tm1, vp=viewport(layout.pos.col=1, layout.pos.row=1, height=5))
#print(tm2, vp=viewport(layout.pos.col=2, layout.pos.row=1, height=5))
#print(tm3, vp=viewport(layout.pos.col=1, layout.pos.row=2, height=5))
#print(legend, vp=viewport(layout.pos.col=2, layout.pos.row=2, height=5))
#how to save this with grid package instead of tmap_arrange?



```
## Inset Map

```{r insetmap}
Londonbb = st_bbox(mapnorm)%>%
    #crs = st_crs()
    st_as_sfc()

?tm_text

tmmain<- tm_shape(mapnorm, bbox=Londonbb)+
  tm_polygons("airbnb_norm",
              style="jenks",
              palette="PuBu")+
  tm_scale_bar(position = c("left", "bottom"), 
               text.size = .60,
               width = 0.20)+
  tm_layout(legend.position =c(0.65, "top"),
            legend.text.size=.65,
            legend.title.size = 1.1,
            frame=FALSE)+
  tm_credits("(c) OpenStreetMap contrbutors and Air b n b", 
             position=c(0.0,0.0))+
  tm_compass(type = "8star", 
             position = c(0.03, 0.1))+ #(0.06, 0.1))+
  #tm_layout(inner.margins=0)
  tm_layout(inner.margin=c(0.04,0.04,0.04,0.4))

inset <- tm_shape(UK_outlinecrop)+
  tm_polygons(col="darkslategray1", lwd=1)+
  tm_shape(Londonbb)+
  tm_borders(col="grey40", lwd = 3)+
  tm_layout(frame=FALSE, bg.color="transparent")+
  tm_shape(Worldcities_extract)+
  tm_symbols(col = "red", scale = .5)+
  tm_text("city_name", xmod=-1.5, ymod=-0.5)

#tmmain
#print(inset,  vp = viewport(0.86, 0.29, width = 0.5, height = 0.55))

tmap_save(tmmain,
          insets_tm = inset,
          insets_vp=viewport(x=0.65, y=0.29, width=.5, height=.55), 
          filename=here::here("R_Maps", "InsetMap.pdf"),
          dpi=600)
```

##Basic Interactive Map

```{r}
tmap_mode("view")

tm_shape(Airbnb) + 
  tm_polygons("Accomodation count", breaks=breaks) 
```



##Advanced Interactive Map






