---
title: "wk6_hw"
author: "Yun Zhao"
date: "11/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
#Wrangling Munging
library(here)
library(janitor)
library(dplyr)
library(stringr)
library(tidyverse)

#Geo
library(sf)
library(sp)
library(rgeos)
library(geojson)
library(geojsonio)
library(raster)
library(fpc)
library(spatstat)
library(geojsonio)

#Plotting
library(maptools)
library(tmap)
library(tmaptools)
library(GISTools)
library(OpenStreetMap)
```

## Load and Preview data
Open Street Map POI data
  Instead: https://www.getthedata.com/open-pubs
  Instead: http://download.slipo.eu/results/osm-to-csv/europe/great-britain/
Police data, Brighton is within the Sussex Police force
LSOA shapefile
```{r}
pubs <- read_csv(here::here("CSVs", "open_pubs.csv"), col_names = FALSE)

#remove lat/long  rows with na's, if any.
pubs$X7 <- as.numeric(pubs$X7)
pubs$X8 <- as.numeric(pubs$X8)

# subset rows that are not na
pubs <- pubs[!is.na(pubs$X7),]
pubs <- pubs[!is.na(pubs$X8),]

pubspts <- SpatialPointsDataFrame(coords = c(pubs[,c("X7", "X8")]),
                                  proj4string = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"),
                                  data = pubs)%>%
  st_as_sf()


# create spatial points from XY  data set # https://riptutorial.com/r/example/6863/create-spatial-points-from-xy-data-set
```
### Study Area
```{r}
lsoa <- st_read(here::here('Shapefiles', "Lower_Super_Output_Area_(LSOA)_IMD2019__(WGS84)-shp",
                           '5878ebcf-f091-4bde-8864-aa547dd45afd2020330-1-8g4usn.8evuq.shp'))


brighton <- lsoa%>%
  filter(str_detect(`lsoa11nm`,'^Brighton and Hove'))
head(brighton)
qtm(brighton)

pubs_brighton <- brighton %>%
  st_join(., y=pubspts, left = TRUE)

pubs_brighton <- pubspts %>%
  st_join(., y=brighton)

crs(pubspts)
crs(brighton)

tmap_mode('plot')
tm_shape(lsoa)+
  tm_borders()
tm_shape(brighton)+
  tm_polygons()+
tm_shape(pubspts)+
  tm_dots()



qtm(pubs_brighton)
head(brighton)
```


```{r}

pubspts.ppp %>%
  density(., sigma=500) %>%
  plot()

crime <- read_csv(here::here('CSVs', '2021-09-sussex-street.csv'))%>%
  clean_names()
                   
```

What could a DBSCAN, Moran’s I analysis of pub locations tell us about where instances of anti-social behaviour might occur?
What limitations might there be in just looking at categorical point data?
What other information might we need to carry out a better analysis?
What are the limitations of spatial pattern analysis?

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
