---
title: "wk7_prac_andy"
author: "Yun Zhao"
date: "11/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(tmap)
library(geojsonio)
library(plotly)
library(rgdal)
library(broom)
library(mapview)
library(crosstalk)
library(sf)
library(sp)
library(spdep)
library(car)
library(fs)
library(janitor)
```

eval=false is so it doesn't download every time you knit
```{r eval=FALSE}
download.file("https://data.london.gov.uk/download/statistical-gis-boundary-files-london/9ba8c833-6370-4b11-abdc-314aa020d5e0/statistical-gis-boundaries-london.zip",
              destfile="Data/statistical-gis-boundaries-london.zip")

```{r}
listfiles<-dir_info(here::here("Data")) %>%
  dplyr::filter(str_detect(path, ".zip")) %>%
  dplyr::select(path)%>%
  pull()%>%
  #print out the .gz file
  print()%>%
  as.character()%>%
  utils::unzip(exdir=here::here("Data"))
```


```{r}
Londonwards<-dir_info(here::here("Data",
                                 "statistical-gis-boundaries-london",
                                 "ESRI"))%>%
  #$ means exact match
  dplyr::filter(str_detect(path,
                           "London_Ward_CityMerged.shp$"))%>%
  dplyr::select(path)%>%
  pull()%>%
  #read in the file in
  st_read()
```


```{r}
qtm(Londonwards)
```

```{r}
LondonWardProfiles <- read_csv("https://data.london.gov.uk/download/ward-profiles-and-atlas/772d2d64-e8c6-46cb-86f9-e52b4c7851bc/ward-profiles-excel-version.csv",
                               na = c("", "NA", "n/a"),
                               locale = locale(encoding = 'Latin1'),
                               col_names = TRUE)
```
```{r}
#check all of the columns have been read in correctly
Datatypelist <- LondonWardProfiles %>%
  summarise_all(class) %>%
  pivot_longer(everything(),
               names_to="All_variables",
               values_to="Variable_class")

Datatypelist
```
```{r}
LonWardProfiles <- Londonwards%>%
  left_join(.,
            LondonWardProfiles,
            by = c("GSS_CODE" = "New code"))

#let's map our dependent variable to see if the join has worked:
tmap_mode("plot")
qtm(LonWardProfiles,
    fill = "Average GCSE capped point scores - 2014",
    borders = NULL,
    fill.palette = "Blues")
```

```{r eval=FALSE}
london_schools <- read_csv("https://data.london.gov.uk/download/london-schools-atlas/57046151-39a0-45d9-8dc0-27ea7fd02de8/all_schools_xy_2016.csv")

```
```{r}
lon_schools_sf <- st_as_sf(london_schools,
                           coords = c("x","y"),
                           crs = 4326)

lond_sec_schools_sf <- lon_schools_sf %>%
  filter(PHASE=="Secondary")

tmap_mode("plot")
qtm(lond_sec_schools_sf)
```
```{r}
q <- qplot(x = `Unauthorised Absence in All Schools (%) - 2013`,
           y = `Average GCSE capped point scores - 2014`,
           data=LonWardProfiles)

#plot with a regression line - note, I've added some jitter here as the x-scale is rounded
q + stat_smooth(method="lm", se=FALSE, size=1) +
  geom_jitter()
```
jitter or would all be on top of each other can't see
the term resideuals means observed? vs actual?
stat_smooth adds the regression line


```{r}
#run the linear regression model and store its outputs in an object called model1
Regressiondata<- LonWardProfiles%>%
  clean_names()%>%
  dplyr::select(average_gcse_capped_point_scores_2014,
                unauthorised_absence_in_all_schools_percent_2013)

#now model
model1 <- Regressiondata %>%
  lm(average_gcse_capped_point_scores_2014 ~
               unauthorised_absence_in_all_schools_percent_2013,
     data=.)
```
Tilda means modeled by
lm means linear model. lm(observed, expected)




# Assumptions

## Linear
variables are normal can help us tell

```{r}
LonWardProfiles <- LonWardProfiles %>%
  clean_names()

ggplot(LonWardProfiles, aes(x=average_gcse_capped_point_scores_2014)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = 5) +
  geom_density(colour="red",
               size=1,
               adjust=1)
```


```{r}
ggplot(LonWardProfiles, aes(x=unauthorised_absence_in_all_schools_percent_2013)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = 0.1) +
  geom_density(colour="red",
               size=1,
               adjust=1)
```

```{r}
ggplot(LonWardProfiles, aes(x=median_house_price_2014)) +
  geom_histogram()
```
```{r}
qplot(x = median_house_price_2014,
      y = average_gcse_capped_point_scores_2014,
      data=LonWardProfiles)
```
```{r}
ggplot(LonWardProfiles, aes(x=log(median_house_price_2014))) +
  geom_histogram()
```

```{r}
symbox(~median_house_price_2014,
       LonWardProfiles,
       na.rm=T,
       powers=seq(-3,3,by=.5))
```

```{r}
ggplot(LonWardProfiles, aes(x=(median_house_price_2014)^-1)) +
  geom_histogram()
```


# Assumption 2
could use tukey's...but not recommended hard to interpret
```{r}
model_data <- model1 %>%
  augment(., Regressiondata)

#plot residuals
model_data%>%
dplyr::select(.resid)%>%
  pull()%>%
  qplot()+
  geom_histogram()
```
```{r}
Regressiondata2<- LonWardProfiles%>%
  clean_names()%>%
  dplyr::select(average_gcse_capped_point_scores_2014,
         unauthorised_absence_in_all_schools_percent_2013,
         median_house_price_2014)

model2 <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 +
               log(median_house_price_2014), data = Regressiondata2)

#show the summary of those outputs
tidy(model2)
```

```{r}
#and for future use, write the residuals out
model_data2 <- model2 %>%
  augment(., Regressiondata2)

# also add them to the shapelayer
LonWardProfiles <- LonWardProfiles %>%
  mutate(model2resids = residuals(model2))
```

```{r}
library(corrr)

Correlation <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(average_gcse_capped_point_scores_2014,
         unauthorised_absence_in_all_schools_percent_2013,
         median_house_price_2014) %>%
  mutate(median_house_price_2014 =log(median_house_price_2014))%>%
    correlate() %>%
  # just focus on GCSE and house prices
  focus(-average_gcse_capped_point_scores_2014, mirror = TRUE)


#visualise the correlation matrix
rplot(Correlation)
```

```{r}
#print some model diagnositcs.
par(mfrow=c(2,2))    #plot to 2 by 2 array
plot(model2)
```

```{r}
tmap_mode("view")
#qtm(LonWardProfiles, fill = "model1_resids")

tm_shape(LonWardProfiles) +
  tm_polygons("model2resids",
              palette = "RdYlBu") +
tm_shape(lond_sec_schools_sf) + tm_dots(col = "TYPE")
```

```{r}
coordsW <- LonWardProfiles%>%
  st_transform(.,crs=27700)%>%
  st_centroid()%>%
  st_geometry()

plot(coordsW)
```

```{r}
LWard_nb <- LonWardProfiles %>%
  poly2nb(., queen=T)

#or nearest neighbours
knn_wards <-coordsW %>%
  knearneigh(., k=4)

LWard_knn <- knn_wards %>%
  knn2nb()

#plot them
plot(LWard_nb, st_geometry(coordsW), col="red")
```

```{r}
Lward.queens_weight <- LWard_nb %>%
  nb2listw(., style="C")

Lward.knn_4_weight <- LWard_knn %>%
  nb2listw(., style="C")
```

```{r}
Queen <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(model2resids)%>%
  pull()%>%
  moran.test(., Lward.queens_weight)%>%
  tidy()

Nearest_neighbour <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(model2resids)%>%
  pull()%>%
  moran.test(., Lward.knn_4_weight)%>%
  tidy()
```

```{r}
library(spatialreg)

slag_dv_model2_queen <- lagsarlm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 +
               log(median_house_price_2014),
               data = LonWardProfiles,
               nb2listw(LWard_nb, style="W"),
               method = "eigen")

#what do the outputs show?
tidy(slag_dv_model2_queen)
```
```{r}
t<-summary(slag_dv_model2_queen)

```

```{r}
#run a spatially-lagged regression model
slag_dv_model2_knn4 <- lagsarlm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 +
               log(median_house_price_2014),
               data = LonWardProfiles,
               nb2listw(LWard_knn,
                        style="W"),
               method = "eigen")

#what do the outputs show?
tidy(slag_dv_model2_knn4)
```
```{r}
LonWardProfiles <- LonWardProfiles %>%
  mutate(slag_dv_model2_knn_resids = residuals(slag_dv_model2_knn4))

KNN4Moran <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(slag_dv_model2_knn_resids)%>%
  pull()%>%
  moran.test(., Lward.knn_4_weight)%>%
  tidy()

KNN4Moran
```
```{r}
sem_model1 <- errorsarlm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 +
               log(median_house_price_2014),
               data = LonWardProfiles,
               nb2listw(LWard_knn, style="C"),
               method = "eigen")

tidy(sem_model1)
```

GWR

```{r}
extradata <- read_csv("https://www.dropbox.com/s/qay9q1jwpffxcqj/LondonAdditionalDataFixed.csv?raw=1")

```
```{r}
LonWardProfiles <- LonWardProfiles%>%
  left_join(.,
            extradata,
            by = c("gss_code" = "Wardcode"))%>%
  clean_names()

#print some of the column names
LonWardProfiles%>%
  names()%>%
  tail(., n=10)
```
```{r}
p <- ggplot(LonWardProfiles,
            aes(x=unauth_absence_schools11,
                y=average_gcse_capped_point_scores_2014))
p + geom_point(aes(colour = inner_outer))
```

```{r}
isitfactor <- LonWardProfiles %>%
  dplyr::select(inner_outer)%>%
  summarise_all(class)

isitfactor
```


```{r}
LonWardProfiles<- LonWardProfiles %>%
  mutate(inner_outer=as.factor(inner_outer))
```

```{r}
model3 <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 +
               log(median_house_price_2014) +
               inner_outer,
             data = LonWardProfiles)

tidy(model3)
```

```{r}
library(spgwr)
```

```{r}
coordsW2 <- st_coordinates(coordsW)

LonWardProfiles2 <- cbind(LonWardProfiles,coordsW2)

GWRbandwidth <- gwr.sel(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 +
                    log(median_house_price_2014) +
                    inner_outer +
                    rate_of_job_seekers_allowance_jsa_claimants_2015 +
                    percent_with_level_4_qualifications_and_above_2011,
                  data = LonWardProfiles2,
                        coords=cbind(LonWardProfiles2$X, LonWardProfiles2$Y),
                  adapt=T)
GWRbandwidth
```


```{r}
gwr.model = gwr(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 +
                    log(median_house_price_2014) +
                    inner_outer +
                    rate_of_job_seekers_allowance_jsa_claimants_2015 +
                    percent_with_level_4_qualifications_and_above_2011,
                  data = LonWardProfiles2,
                coords=cbind(LonWardProfiles2$X, LonWardProfiles2$Y),
                adapt=GWRbandwidth,
                hatmatrix=TRUE,
                se.fit=TRUE)

#print the results of the model
gwr.model
```

```{r}
results <- as.data.frame(gwr.model$SDF)

```


```{r}
LonWardProfiles2 <- LonWardProfiles %>%
  mutate(coefUnauthAbs = results$unauthorised_absence_in_all_schools_percent_2013,
         coefHousePrice = results$log.median_house_price_2014.,
         coefJSA = rate_of_job_seekers_allowance_jsa_claimants_2015,
         coefLev4Qual = percent_with_level_4_qualifications_and_above_2011)
```

```{r}
sigTest = abs(gwr.model$SDF$"log.median_house_price_2014.")-(2 * gwr.model$SDF$"log.median_house_price_2014._se")

```




